Model Search in Relational Domains

### 2016-10-20 algorithm.tex
comments on example
learning

\section{Abstract Conflict Driven Clause Learning}
%
\begin{algorithm2e}[t]
\DontPrintSemicolon
\SetKwFunction{decide}{Decide}
\SetKwFunction{deduce}{deduce}
\SetKwFunction{backtrack}{backtrack}
\SetKwFunction{print}{print}
\SetKwFunction{return}{return}
\SetKwFunction{nondet}{nondet}
\SetKwFunction{run}{run}
\SetKwFunction{learn}{learn}
\SetKwFunction{continue}{continue}
\SetKwFunction{assign}{assign}
\SetKwRepeat{Do}{do}{while}
%\SetKwFunction{assume}{assume}
%\SetKwFunction{isf}{isFeasible}
\SetKwData{conflict}{CONFLICT}
\SetKwData{safe}{safe}
\SetKwData{sat}{SAT}
\SetKwData{unsafe}{unsafe}
\SetKwData{unknown}{unknown}
\SetKwData{true}{true}
\SetKwInOut{Input}{input}
\SetKwInOut{Output}{output}
\SetKwFor{Loop}{Loop}{}{}
\SetKw{KwNot}{not}
\begin{small}
\Input{Set $S$ of abstract deduction transformer, 
$\widehat{ded_{\varphi}}$ for a formula $\varphi$ over an abstract domain $A$}
\Output{The status (\safe or \unsafe) and a counterexample if \unsafe}
$\mathcal{T} \leftarrow \langle\rangle$ \;
$\mathcal{R} \leftarrow \emptyset$ \;
$\mathit{result} \leftarrow \deduce(S,\mathcal{T},\mathcal{R})$ \;
\lIf{$\mathit{result}$ = \conflict} {
  \return \safe}
\While{$true$} 
{
\lIf{$\mathit{result}$ = \sat} {
  \return \unsafe}
  $d \leftarrow decide(\mathit{abs}(\mathcal{T}))$ \;
%  $v \leftarrow v \meet d$ \; 
  $\mathcal{T} \leftarrow \mathcal{T} \cdot d$ \; 
  $\mathcal{R}[|\mathcal{T}|] \leftarrow $\textsf{nil} \;
  $\mathit{result} \leftarrow \deduce(S,v,\mathcal{T},\mathcal{R})$\;
  \Do{$\mathit{result} = \conflict$} {
%    learn() \;
    \lIf{$\neg \mathit{conflict\_analysis}(S,\mathcal{T},\mathcal{R})$} {
      \return \safe
    }
    $\mathit{result} \leftarrow \deduce(S,\mathcal{T},\mathcal{R})$ \;
  }
}
\end{small}
\caption{Abstract Conflict Driven Clause Learning $\langle ACDCL(S) \rangle$ \label{Alg:acdcl}}
\end{algorithm2e}
%
%\subsection{Algorithm for Abstract Conflict Driven Clause Learning}
Algorithm~\ref{Alg:acdcl} presents an overview of the ACDCL algorithm.
The algorithm begins with a set $S$ of abstract deduction transformer 
$\widehat{ded_{\varphi}}$ \pscmt{define in terms of P and A} for a formula $\varphi$ over an abstract 
domain $A$, an abstract valuation $\langle v \leftarrow \top \rangle$, 
and an empty propagation trail and reason trail, $\mathcal{T}$, $\mathcal{R}$, 
respectively.  The propagation trail stores all meet irreducibles inferred 
by the abstract model search phase (deductions and decisions), while the 
reason trail maps elements of propagation trail to transformers that were 
used to derive them. We convert the propagation trail into an abstract value as follows; $\mathit{abs}(\mathcal{T})=\bigsqcap_{m\text{ in }\mathcal{T}}m$ with $\mathit{abs}(\mathcal{T})=\top$ if $\mathcal{T}$ is the empty sequence $\langle\rangle$.

@@

\pscmt{define ``consistent with trail''}

>>>>

\section{Abstract Conflict Driven Clause Learning}
%
\begin{algorithm2e}[t]
\DontPrintSemicolon
\SetKwFunction{decide}{Decide}
\SetKwFunction{deduce}{deduce}
\SetKwFunction{backtrack}{backtrack}
\SetKwFunction{print}{print}
\SetKwFunction{return}{return}
\SetKwFunction{nondet}{nondet}
\SetKwFunction{run}{run}
\SetKwFunction{learn}{learn}
\SetKwFunction{continue}{continue}
\SetKwFunction{assign}{assign}
\SetKwRepeat{Do}{do}{while}
%\SetKwFunction{assume}{assume}
%\SetKwFunction{isf}{isFeasible}
\SetKwData{conflict}{CONFLICT}
\SetKwData{safe}{safe}
\SetKwData{sat}{SAT}
\SetKwData{unsafe}{unsafe}
\SetKwData{unknown}{unknown}
\SetKwData{true}{true}
\SetKwInOut{Input}{input}
\SetKwInOut{Output}{output}
\SetKwFor{Loop}{Loop}{}{}
\SetKw{KwNot}{not}
\begin{small}
\Input{Set $S$ of abstract deduction transformer, 
$\widehat{ded_{\varphi}}$ for a formula $\varphi$ over an abstract domain $A$}
\Output{The status (\safe or \unsafe) and a counterexample if \unsafe}
$\mathcal{T} \leftarrow \langle\rangle$ \;
$\mathcal{R} \leftarrow \emptyset$ \;
$\mathit{result} \leftarrow \deduce(S,\mathcal{T},\mathcal{R})$ \;
\lIf{$\mathit{result}$ = \conflict} {
  \return \safe}
\While{$true$} 
{
\lIf{$\mathit{result}$ = \sat} {
  \return \unsafe}
  $d \leftarrow decide(\mathit{abs}(\mathcal{T}))$ \;
%  $v \leftarrow v \meet d$ \; 
  $\mathcal{T} \leftarrow \mathcal{T} \cdot d$ \; 
  $\mathcal{R}[|\mathcal{T}|] \leftarrow $\textsf{nil} \;
  $\mathit{result} \leftarrow \deduce(S,\mathcal{T},\mathcal{R})$\;
  \Do{$\mathit{result} = \conflict$} {
%    learn() \;
    \lIf{$\neg \mathit{conflict\_analysis}(S,\mathcal{T},\mathcal{R})$} {
      \return \safe
    }
    $\mathit{result} \leftarrow \deduce(S,\mathcal{T},\mathcal{R})$ \;
  }
}
\end{small}
\caption{Abstract Conflict Driven Clause Learning $\langle ACDCL(S) \rangle$ \label{Alg:acdcl}}
\end{algorithm2e}
%
%\subsection{Algorithm for Abstract Conflict Driven Clause Learning}
Algorithm~\ref{Alg:acdcl} presents an overview of the ACDCL algorithm.
The algorithm begins with a set $S$ of abstract deduction transformer 
$\widehat{ded_{\varphi}}$ \pscmt{define in terms of P and A} for a formula $\varphi$ over an abstract 
domain $A$, an abstract valuation $\langle v \leftarrow \top \rangle$, 
and an empty propagation trail and reason trail, $\mathcal{T}$, $\mathcal{R}$, 
respectively.  The propagation trail stores all meet irreducibles inferred 
by the abstract model search phase (deductions and decisions), while the 
reason trail maps elements of propagation trail to transformers that were 
used to derive them. A propagation trail is mapped to an abstract value as 
follows; $\mathit{abs}(\mathcal{T})=\bigsqcap_{m \in \mathcal{T}}m$ with 
$\mathit{abs}(\mathcal{T})=\top$ if $\mathcal{T}$ is the empty sequence $\langle\rangle$.

@@

%\pscmt{define ``consistent with trail''}

### 2016-10-20 learning.tex
learning

\paragraph {\textbf{Lifting UIP to relational domains}}
An UIP is a special node in the abstract conflict graph such that 
any path from the last decision node to the conflict node must pass 
through it.  Unlike a SAT solver, a UIP for relational domain may 
involve meet irreducibles that contains the same variables.  The 
first-UIP is a unique node closest to the conflict, and the last UIP 
is the decision node itself.  Computing UIPs ensures asserting
cuts~\cite{cdcl,DBLP:journals/fmsd/BrainDGHK14}, that is, it 
yields clauses that generates new deduction after backtracking.  
Every cut in the graph is a reason for conflict that can be 
used in learning.  An abstract first-UIP algorithm~\cite{DBLP:journals/fmsd/BrainDGHK14} 
traverses the trail $\mathcal{T}$ starting from the conflict node and 
computes a cut that suffices to produce a conflict. 
\Omit {
contradict with the transformer stored in $\mathcal{R}[\bot]$.  
}
For our example, there exists multiple incomaparable reasons for conflict,
marked as {\em cut0, cut1}, in figure~\ref{conflict}.  Here, cut0 is the first UIP.  
Choosing cut0 yields a learnt clause 
$(x+y<15 \vee -x-y<-15 \vee -x+y<5 \vee x-y<-5 \vee y-10<0 \vee -y+10<0)$, 
which is obtained by negating the reason for conflict.  
%
\begin{figure}[t]
%\caption{An example of Octagon}\label{octagon}
\scalebox{.65}{\import{figures/}{conflict_graph.pspdftex}}
\caption{\label{conflict} Abstract Conflict Graph}
\end{figure} 
%    
\paragraph{\textbf{Learning in Relational Lattice}}
Learning in propositional solver yields a asserting clause~\cite{cdcl} 
that express the negation of the conflict reasons.  However, we model 
learning in relational domains as learning a transformer which infers 
new meet irreducibles using the abstract unit rule called 
{\em abstract unit transformer (AUnit)}.  We add $AUnit$ to the set of 
transformers \rmcmt{give symbol} used during model search. $AUnit$ is 
a generalization of propositional unit rule for numerical domains.  For 
an abstract lattice $A$ with complementable meet irreducibles and a set 
of meet irreducibles $C \subseteq A$ of $\varphi$ such that $\bigsqcap C$ does not 
satisfy $\varphi$, $AUnit_C: A \rightarrow A$ is formally defined as follows. 
\[ AUnit(a) =
 \begin{cases}
  \bot       & \quad \text{if } a \sqsubseteq \meet{C} \qquad \qquad \qquad
  \qquad \qquad \qquad \quad (1)\\
  \bar{t}    & \quad \text{if } t \in C \; \text{and} \; \forall t' \in C
  \setminus \{t\}. a  \sqsubseteq t' \qquad \quad (2) \\
  \bar{t}    & \quad \text{if } t \in C \; \text{and} \; \forall t' \in C \setminus \{t\}. a
  \not\sqsubseteq t', \; \text{then} \\ 
             & \quad \forall a' \subseteq decomp(a) \;  \text{and} \; \forall a'' \in a \setminus \{a'\}, \\ 
             & \quad \meet{a''} \sqsubseteq r' \; \text{and} \; a'
             \not\sqsubseteq t   \qquad \qquad \qquad \qquad (3) \\
  \top & \quad \text{otherwise} \qquad \qquad \qquad \qquad \qquad \qquad \quad (4) \\
 \end{cases}
\]

\pscmt{explain which cases are different in non-relational domains}
For example, let $C = \langle x+z \geq 10,  w \geq 0 \rangle$ 
be a conflicting element that does not satisfy $\varphi$ and let $a= \langle x+y \geq 2, y+z \geq 5,
x+z \leq 15, w \geq 1, w \leq 1 \rangle$.  Applying the rule (3) of $AUnit_C$, we get a 
decomposition of $a$ into $a'$ and $a''$ where 
$a'= \langle x+y \geq 2, y+z \geq 5, x+z \leq 15 \rangle \not\sqsubseteq (x+z \geq 10)$, and 
$a''= \langle w \geq 1, w \leq 1 \rangle \sqsubseteq (w \geq 0)$.  Then, $AUnit_C(a) = (x+z < 10)$, 
which constrains the bound of the octagonal constraint.  
Let us consider another example, where $C = \langle x \geq 2, x \leq 5, y \leq 7 \rangle$
and let $a = \langle x \geq 3, x \leq 4, y \geq 5, y \leq 6 \rangle$, then $AUnit_C(a) = \bot$ using
the first rule, since $a \sqsubseteq C$.  Now, let $a' = \langle x \geq 3, x \leq 4 \rangle$, 
then $AUnit_C(a') = (y > 7)$ using second rule, since $a' \sqsubseteq (x \geq 2)$
and $a' \sqsubseteq (x \leq 5)$.

@@

The abstract clause learning and backjumping procedures in the abstract 
conflict graph is stated in terms of the state of ACDCL solver as follows. 
\[AbsLearn: \quad  (\mathcal{E},S) \rightarrow (\mathcal{E},S \wedge
\mathcal{L}) \quad \text{if} \; \mathcal{L} \notin S \; \textrm{and}
\; (S \wedge \mathcal{L}) \; \text{is not UNSAT} \]
\[AbsBackjump: \quad (\mathcal{E}_1(m,s)\mathcal{E}_2,S) \rightarrow
(\mathcal{E}_1,S) \quad \text{if} \; (\mathcal{E}_1,S) \; \text{is
consistent} \]   
\pscmt{What is $\mathcal{E}$? What is $m$? What is the notation $\mathcal{E}_1(m,s)\mathcal{E}_2$? What does consistent mean?}
\pscmt{Is there any difference between non-relational and relational?}

>>>>

\paragraph {\textbf{Lifting UIP to relational domains}}
An UIP is a special node in the abstract conflict graph such that 
any path from the last decision node to the conflict node must pass 
through it.  Unlike a SAT solver, a UIP for relational domain may 
involve meet irreducibles that contains the same variables.  The 
first-UIP is a unique node closest to the conflict, and the last UIP 
is the decision node itself.  Computing UIPs ensures asserting
cuts~\cite{cdcl,DBLP:journals/fmsd/BrainDGHK14}, that is, it 
yields clauses that generates new deduction after backtracking.  
Every cut in the graph is a reason for conflict that can be 
used in learning.  An abstract first-UIP algorithm~\cite{DBLP:journals/fmsd/BrainDGHK14} 
traverses the trail $\mathcal{T}$ starting from the conflict node and 
computes a cut that suffices to produce a conflict. 
\Omit {
contradict with the transformer stored in $\mathcal{R}[\bot]$.  
}
For our example, there exists multiple incomaparable reasons for conflict,
marked as {\em cut0, cut1}, in figure~\ref{conflict}.  Here, cut0 is the first UIP.  
Choosing cut0 yields a learnt clause 
$(x+y<15 \vee -x-y<-15 \vee -x+y<5 \vee x-y<-5 \vee y-10<0 \vee -y+10<0)$, 
which is obtained by negating the reason for conflict.  
%
\begin{figure}[t]
%\caption{An example of Octagon}\label{octagon}
\scalebox{.65}{\import{figures/}{conflict_graph.pspdftex}}
\caption{\label{conflict} Abstract Conflict Graph}
\end{figure} 
%    
\paragraph{\textbf{Learning in Relational Lattice}}
Learning in propositional solver yields a asserting clause~\cite{cdcl} 
that express the negation of the conflict reasons.  However, we model 
learning in relational domains as learning a transformer which infers 
new meet irreducibles using the abstract unit rule called 
{\em abstract unit transformer (AUnit)}.  We add $AUnit$ to the set of 
transformers \rmcmt{give symbol} used during model search. $AUnit$ is 
a generalization of propositional unit rule for numerical domains.  For 
an abstract lattice $A$ with complementable meet irreducibles and a set 
of meet irreducibles $C \subseteq A$ of $\varphi$ such that $\bigsqcap C$ does not 
satisfy $\varphi$, $AUnit_C: A \rightarrow A$ is formally defined as follows. 
\[ AUnit(a) =
 \begin{cases}
  \bot       & \quad \text{if } a \sqsubseteq \meet{C} \qquad \qquad \qquad
  \qquad \qquad \qquad \quad (1)\\
  \bar{t}    & \quad \text{if } t \in C \; \text{and} \; \forall t' \in C
  \setminus \{t\}. a  \sqsubseteq t' \qquad \quad \; \; (2) \\
  \bar{t}    & \quad \text{if } t \in C \; \text{and} \; \forall t' \in C \setminus \{t\}. a
  \not\sqsubseteq t', \; \text{then} \\ 
             & \quad \forall a' \subseteq decomp(a) \;  \text{and} \; \forall a'' \in a \setminus \{a'\}, \\ 
             & \quad \meet{a''} \sqsubseteq r' \; \text{and} \; a'
             \not\sqsubseteq t   \qquad \qquad \qquad \qquad \; \; (3) \\
  \top & \quad \text{otherwise} \qquad \qquad \qquad \qquad \qquad \qquad \quad (4) \\
 \end{cases}
\]
$AUnit$ returns $\bot$ if $C$ is conflicting following rule (1).  Rule (2) and
(3) of $AUnit$ infers a valid meet irreducible, which implies that $C$ is unit. 
However, rule (3) is domain-specific and only applies to numerical abstract
domains.  For example, let $C = \langle x+z \geq 10,  w \geq 0 \rangle$ 
be a conflicting element that does not satisfy $\varphi$ and let $a= \langle x+y \geq 2, y+z \geq 5,
x+z \leq 15, w \geq 1, w \leq 1 \rangle$.  Applying rule (3) of $AUnit_C$, we get a 
decomposition of $a$ into $a'$ and $a''$ where 
$a'= \langle x+y \geq 2, y+z \geq 5, x+z \leq 15 \rangle \not\sqsubseteq (x+z \geq 10)$, and 
$a''= \langle w \geq 1, w \leq 1 \rangle \sqsubseteq (w \geq 0)$.  Then, $AUnit_C(a) = (x+z < 10)$, 
which constrains the bound of the octagonal constraint.  
Let us consider another example, where $C = \langle x \geq 2, x \leq 5, y \leq 7 \rangle$
and let $a = \langle x \geq 3, x \leq 4, y \geq 5, y \leq 6 \rangle$, then $AUnit_C(a) = \bot$ using
the rule (1), since $a \sqsubseteq C$.  Now, let $a' = \langle x \geq 3, x \leq 4 \rangle$, 
then $AUnit_C(a') = (y > 7)$ using rule (2), since $a' \sqsubseteq (x \geq 2)$
and $a' \sqsubseteq (x \leq 5)$.  However, if none of the above rule holds true, $AUnit$ 
returns $\top$.

@@

The abstract clause learning and backjumping procedures in the abstract 
conflict graph is stated in terms of the state of ACDCL solver as follows. 
\[AbsLearn: \quad  (\mathcal{E},S) \rightarrow (\mathcal{E},S \wedge
\mathcal{L}) \quad \text{if} \; \mathcal{L} \notin S \; \textrm{and}
\; (S \wedge \mathcal{L}) \; \text{is not UNSAT} \]
\[AbsBackjump: \quad (\mathcal{E}_1(m,s)\mathcal{E}_2,S) \rightarrow
(\mathcal{E}_1,S) \quad \text{if} \; (\mathcal{E}_1,S) \; \text{is not in conflict}} \]
